\declchap{Technologies in Use}{techno}

The C++ programming language, developed through out this \MT, is based on two technologies, which are 
introduced in this chapter. The first technology is the \jbmps\ language engineering environment, which 
provides core foundations and means for incremental language construction. The second technology is
the \mbdp. These technologies are discussed in this chapter.

\input{newchapters/MPS}

\input{newchapters/MBEDDR}



\declsec{C++ Language}{cpplangitself}

This work is basically an implementation of the \cpppl\ in \jbmps. No matter, the whole work 
is about the C++ implementation, it is infeasible to describe in detail the language 
here. Thus, we only give references to literature about the language in this chapter. Certain 
aspects of C++ are explained in depth during the description of the C++ implementation in \jbmps\ itself in 
the \rchap{cpp}.

Various literature is available, to get acquainted with the language closer. The most complete
guide to the language, covering \rg{stl} as well is \cite{stroustrupcpp2000}. A more 
easy-to-read and more suitable for beginners book on C++ is \cite{schildtcpp}. A newer book,
oriented towards intermediate level C++ programmers and updated to the recent C++ 11 standard (referenced as \cite{cpp11}) and describing 
\rg{stl} as well is \cite{pratacpp}.

A collection of techniques to get more effective C++ programs is gathered in \cite{meyerseffcpp}. Templates are explained in 
detail in \cite{josuttistemplates}. A book dedicated to ad hoc polymorphism and advanced template meta programming can be 
also of interest as an approach to language engineering on top of C++ in a certain sense, \cite{alexandrescu}.

The \cpppl\ is a mature language, with long traditions, and high flexibility, \cite{alexandrescu} can serve as an example. 
It would not be possible to simplify the language, removing features from it, which would restrict the language use, 
most importantly the \rg{stl} support. Thus in this work we try to research, how the editor can be more supportive 
for a user of the \emph{existing} complex language, how to eliminate usual mistakes made while programming, as 
well as how to provide help in a code structuring.